---
title: "Task 6: Virtual Screening and Hit Expansion"
author: "Pharmacoinformatics BI-424 | AChE Inhibitor Discovery"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    number_sections: true
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width  = 10,
  fig.height = 6,
  dpi = 150)
```

------------------------------------------------------------------------

# Install & Load Packages

```{r packages}
required_packages <- c(
  "tidyverse",
  "ggrepel",
  "patchwork",
  "viridis",
  "knitr",
  "kableExtra",
  "scales",
  "RColorBrewer",
  "ggridges"
)

for (pkg in required_packages) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg, repos = "https://cloud.r-project.org")
    library(pkg, character.only = TRUE)
  }
}

cat("All packages installed and loaded.\n")
```

------------------------------------------------------------------------

# Data Loading

**Files needed in this folder:** - `ACHE_QSAR_dataset.csv` -
`potent_ache_inhibitors.csv` (ChEMBL export)

```{r load-data}
# -- Load the full curated QSAR dataset -----------------------------
main_df <- read.csv("ACHE_QSAR_dataset.csv", stringsAsFactors = FALSE)

# Standardise column names
colnames(main_df)[colnames(main_df) == "pChEMBL.Value"]    <- "pIC50"
colnames(main_df)[colnames(main_df) == "Molecular_Weight"] <- "MW"

# Handle duplicate AlogP column 
if (sum(colnames(main_df) == "AlogP") > 1) {
  dup_idx <- which(colnames(main_df) == "AlogP")
  main_df <- main_df[, -dup_idx[-1]]
}
colnames(main_df)[colnames(main_df) == "AlogP"] <- "logP"

# -- Load potent inhibitor ChEMBL export ---------------------
raw_potent <- read.csv("potent_ache_inhibitors.csv", stringsAsFactors = FALSE)

# Extract and standardise relevant columns
potent_df <- raw_potent %>%
  select(
    ChEMBL_ID      = `Molecule.ChEMBL.ID`,
    MW              = `Molecular.Weight`,
    RO5_Violations  = `X.RO5.Violations`,
    logP            = AlogP,
    SMILES          = Smiles,
    IC50_nM         = `Standard.Value`,
    pIC50           = `pChEMBL.Value`,
    Potency_Cat     = `Potency.Category`,
    HBD             = `Ligand.Efficiency.BEI`,   # placeholder
    LE              = `Ligand.Efficiency.LE`,
    LLE             = `Ligand.Efficiency.LLE`
  ) %>%
  mutate(
    pIC50          = as.numeric(pIC50),
    MW             = as.numeric(MW),
    logP           = as.numeric(logP),
    RO5_Violations = as.numeric(RO5_Violations),
    IC50_nM        = as.numeric(IC50_nM),
    LE             = as.numeric(LE),
    LLE            = as.numeric(LLE)
  ) %>%
  filter(!is.na(pIC50), !is.na(SMILES), SMILES != "",
         !is.na(MW), !is.na(logP)) %>%
  distinct(ChEMBL_ID, .keep_all = TRUE)

cat("=== Main QSAR Dataset ===\n")
cat("Compounds:", nrow(main_df), "\n")
cat("pIC50 range:", round(min(as.numeric(main_df$pIC50), na.rm=TRUE),2),
    "to", round(max(as.numeric(main_df$pIC50), na.rm=TRUE),2), "\n\n")

cat("=== Potent Inhibitors (ChEMBL Export) ===\n")
cat("Compounds:", nrow(potent_df), "\n")
cat("pIC50 range:", round(min(potent_df$pIC50, na.rm=TRUE),2),
    "to", round(max(potent_df$pIC50, na.rm=TRUE),2), "\n")
```

------------------------------------------------------------------------

# Virtual Screening Strategy

## Rationale

Virtual screening (VS) is used to prioritise a small number of compounds
from a larger pool for the computationally expensive docking step (max
10 compounds per project constraints). Our strategy is **QSAR-based
retrospective screening**: we screen from the ChEMBL-validated AChE
inhibitor dataset, applying sequential property filters followed by a
multi-parameter composite scoring function that integrates potency,
ligand efficiency, drug-likeness, and lipophilicity optimality.

This approach is justified because: 1. All compounds have experimentally
validated AChE inhibitory activity 2. pIC50 values are standardised from
IC50 measurements 3. Drug-likeness filtering ensures candidates have
realistic pharmacokinetic potential 4. Ligand efficiency normalises for
molecular size -- preventing selection of large, potent-but-inefficient
molecules 5. Composite scoring is applied before any diversity filter;
in a production pipeline, Tanimoto-based clustering would follow to
ensure scaffold diversity among the final docking candidates

## Screening Library Source

The starting pool consists of all **potent AChE inhibitors** (pIC50 \>=
7.0, i.e. IC50 \<= 100 nM) from the ChEMBL CHEMBL220 dataset

------------------------------------------------------------------------

# Step 1: Drug-likeness Filtering (Lipinski Ro5)

## Filter Criteria

| Property | Threshold | Rationale |
|------------------------|------------------------|------------------------|
| Molecular Weight (MW) | \<= 500 Da | Lipinski Ro5: predicts oral absorption |
| logP | \<= 5.0 | Lipinski Ro5: limits passive diffusion through membranes |
| RO5 Violations | \<= 1 | Allows one exception (Veber rule -- some CNS drugs violate) |
| MW minimum | \>= 150 Da | Removes fragments too small for meaningful binding |
| logP minimum | \>= -2.0 | Removes overly hydrophilic compounds with poor CNS penetration |

```{r ro5-filter}
# -- Record sizes at each step for the funnel ----------------------------------
n_start <- nrow(potent_df)
cat("Starting pool (pIC50 >= 7.0):", n_start, "compounds\n\n")

# Step 1a: Ro5 -- MW, logP, violations
after_ro5 <- potent_df %>%
  filter(MW <= 500, logP <= 5.0, RO5_Violations <= 1)
cat("After Ro5 (MW<=500, logP<=5, viol<=1):", nrow(after_ro5), "\n")

# Step 1b: MW minimum
after_mw_min <- after_ro5 %>% filter(MW >= 150)
cat("After MW >= 150:                      ", nrow(after_mw_min), "\n")

# Step 1c: logP range
after_logp <- after_mw_min %>% filter(logP >= -2.0)
cat("After logP >= -2.0:                   ", nrow(after_logp), "\n")

# Step 1d: Deduplicate
screen_library <- after_logp %>% distinct(ChEMBL_ID, .keep_all = TRUE)
cat("After deduplication:                  ", nrow(screen_library), "\n")
cat("\nFinal screening library size:", nrow(screen_library), "compounds\n")
```

## Table T6-1: Filtering Summary

```{r filter-table}
filter_summary <- tibble(
  Step = 1:5,
  Filter_Applied = c(
    "Starting pool: pIC50 >= 7.0 (potent AChE inhibitors)",
    "Lipinski Ro5: MW <= 500 Da, logP <= 5.0, RO5 violations <= 1",
    "Fragment removal: MW >= 150 Da",
    "Hydrophilicity filter: logP >= -2.0",
    "Deduplication by ChEMBL ID"
  ),
  N_Remaining = c(
    n_start,
    nrow(after_ro5),
    nrow(after_mw_min),
    nrow(after_logp),
    nrow(screen_library)
  ),
  N_Removed = c(
    0,
    n_start - nrow(after_ro5),
    nrow(after_ro5) - nrow(after_mw_min),
    nrow(after_mw_min) - nrow(after_logp),
    nrow(after_logp) - nrow(screen_library)
  ),
  Pct_Remaining = paste0(round(100 * c(
    n_start, nrow(after_ro5), nrow(after_mw_min),
    nrow(after_logp), nrow(screen_library)
  ) / n_start, 1), "%")
)

kable(filter_summary,
      caption = "Table T6-1: Virtual Screening Library Construction -- Sequential Filters") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width=TRUE) %>%
  column_spec(3, bold=TRUE, color="#1a6e3d") %>%
  column_spec(4, color="#c62828") %>%
  row_spec(nrow(filter_summary), bold=TRUE, background="#fff9c4")
```

------------------------------------------------------------------------

# Step 2: Composite Scoring for QSAR-Based Ranking

## Scoring Function Design

Each compound in the filtered library receives a **composite score**
integrating four independent properties. This is a multi-parameter
optimisation (MPO) approach standard in drug discovery for balancing
potency against drug-likeness.

$$CompositeScore = (0.50 \times pIC50_{norm}) + (0.25 \times LE_{norm}) + (0.15 \times S_{DL}) + (0.10 \times S_{logP})$$
Where: - ${\text{pIC50}_{norm}}$ = min-max normalised pIC50 (primary
potency) - ${\text{LE}_{norm}}$ = min-max normalised ligand efficiency
(pIC50 per heavy atom) - $S_{\text{DL}}$ = drug-likeness score (penalty
for Ro5 violations) - $S_{\text{logP}}$ = lipophilicity optimality score
(penalty for logP \> 3.5)

**Weights rationale:** - pIC50 (50%): Primary measure -- biological
activity is the most critical criterion - LE (25%): Prevents selection
of large/bloated molecules that are potent only due to size -
Drug-likeness (15%): Ensures oral bioavailability potential - logP
(10%): Excessive lipophilicity (ClogP \> 3.5) is associated with poor
solubility, hERG binding risk, and metabolic instability

```{r composite-scoring}
# -- Compute composite score ----------------------------------------------------
ranked_library <- screen_library %>%
  mutate(
    # Heavy atom count estimate (MW / average heavy atom weight ~13.8)
    Heavy_Atoms = pmax(1, round(MW / 13.8)),

    # Ligand efficiency: pIC50 per heavy atom
    LE_calc = round(pIC50 / Heavy_Atoms, 4),

    # Min-max normalise pIC50
    pIC50_norm = (pIC50 - min(pIC50, na.rm=TRUE)) /
                 (max(pIC50, na.rm=TRUE) - min(pIC50, na.rm=TRUE)),

    # Min-max normalise LE
    LE_norm    = (LE_calc - min(LE_calc, na.rm=TRUE)) /
                 (max(LE_calc, na.rm=TRUE) - min(LE_calc, na.rm=TRUE)),

    # Drug-likeness score
    DL_score   = case_when(
      RO5_Violations == 0 ~ 1.00,
      RO5_Violations == 1 ~ 0.85,
      TRUE                ~ 0.60
    ),

    # logP optimality score (AChE is a CNS target -- logP 1-3.5 is ideal)
    logP_score = case_when(
      logP >= 1.0 & logP <= 3.5 ~ 1.00,
      logP >= 0.0 & logP <= 5.0 ~ 0.85,
      logP >= -2  & logP <  0.0 ~ 0.70,
      TRUE                       ~ 0.60
    ),

    # Composite score (weighted sum)
    Composite_Score = round(
      0.50 * pIC50_norm +
      0.25 * LE_norm    +
      0.15 * DL_score   +
      0.10 * logP_score,
      4
    )
  ) %>%
  arrange(desc(Composite_Score)) %>%
  mutate(Rank = row_number())

cat("Ranked library size:", nrow(ranked_library), "compounds\n")
cat("Score range:", round(min(ranked_library$Composite_Score),4),
    "to", round(max(ranked_library$Composite_Score),4), "\n")
cat("\nTop 5 scores:\n")
ranked_library %>%
  select(Rank, ChEMBL_ID, pIC50, LE_calc, Composite_Score) %>%
  head(5) %>%
  print()
```

------------------------------------------------------------------------

# Step 3: Top 10 Selection for Molecular Docking

## Selection Criteria

The top 10 compounds by composite score are selected for molecular
docking (maximum 10 per project constraints). To promote chemical
diversity among the selected candidates, we verify that the top 10
include representatives from different potency categories.

```{r top10-selection}
top10 <- ranked_library %>%
  head(10) %>%
  mutate(
    pIC50    = round(pIC50, 2),
    MW       = round(MW, 1),
    logP     = round(logP, 2),
    LE_calc  = round(LE_calc, 3),
    Composite_Score = round(Composite_Score, 4)
  )

cat("=== TOP 10 COMPOUNDS SELECTED FOR DOCKING ===\n\n")
for (i in 1:10) {
  cat(sprintf(
    "Rank %2d | %-20s | pIC50: %5.2f | IC50: %8.3f nM | MW: %5.1f | logP: %4.2f | LE: %.3f | Score: %.4f\n",
    i, top10$ChEMBL_ID[i], top10$pIC50[i], top10$IC50_nM[i],
    top10$MW[i], top10$logP[i], top10$LE_calc[i], top10$Composite_Score[i]
  ))
}
```

## Table T6-2: Top 10 Docking Candidates

```{r top10-table}
top10_display <- top10 %>%
  select(Rank, ChEMBL_ID, pIC50, IC50_nM, MW, logP, RO5_Violations,
         LE_calc, Composite_Score, Potency_Cat)

kable(top10_display,
      col.names = c("Rank", "ChEMBL ID", "pIC50", "IC50 (nM)",
                    "MW (Da)", "logP", "RO5 Viol.", "LE", "Composite Score", "Potency Category"),
      caption = "Table T6-2: Top 10 Compounds Selected for Molecular Docking") %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width=TRUE) %>%
  row_spec(1:3, background="#c8e6c9", bold=TRUE) %>%
  row_spec(4:7, background="#e8f5e9") %>%
  column_spec(3, bold=TRUE) %>%
  column_spec(9, bold=TRUE, color="#1a6e3d")
```

## Diversity Note and Limitation

The top 10 compounds by composite score include five structural
analogues from the same methoxyphenyl-sulfamide scaffold series
(CHEMBL3585780-84), differing only in the position and number of methoxy
substituents. These share identical pIC50 values (10.57) and were ranked
together due to equal composite scores.

In a production virtual screening pipeline, a Tanimoto similarity filter
(cutoff \< 0.7 on Morgan fingerprints) or maximum common substructure
clustering would be applied at this stage to ensure scaffold diversity
among docking candidates. Given the project constraints and
computational scope, all top 10 compounds are retained for docking.
Ranking 6-10 introduce a second distinct scaffold
(tetrahydroisoquinolinone with aryl amide substituents), providing some
chemical diversity across the docking set.

Additionally, the pIC50 values of 10.57 (IC50 = 0.027 nM) for the top
compounds represent sub-picomolar potency -- exceptionally rare in
medicinal chemistry. These data originate from a single ChEMBL
publication and may reflect a specific assay format. This is noted as a
data quality caveat; the composite scoring and filtering methodology
remain valid regardless of the absolute potency values.

------------------------------------------------------------------------

## Table T6-3: SMILES Strings for Docking Input

```{r smiles-table}
smiles_table <- top10 %>%
  mutate(Docking_ID = paste0("ACHE_Cand_", sprintf("%02d", Rank), "_", ChEMBL_ID)) %>%
  select(Rank, ChEMBL_ID, Docking_ID, SMILES, pIC50)

kable(smiles_table,
      caption = "Table T6-3: SMILES Strings for Top 10 Docking Candidates") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width=TRUE) %>%
  column_spec(4, width= "45%", extra_css = "font-size: 9px;")
```

------------------------------------------------------------------------

# Figure T6-1: Virtual Screening Funnel

```{r fig-funnel, fig.height=5}
funnel_data <- tibble(
  Stage = factor(c(
    paste0("Full QSAR\nDataset\nn=", nrow(main_df)),
    paste0("Potent Pool\n(pIC50>=7.0)\nn=", n_start),
    paste0("After Ro5\nFilter\nn=", nrow(after_ro5)),
    paste0("After MW/logP\nCleaning\nn=", nrow(after_logp)),
    paste0("Final Screening\nLibrary\nn=", nrow(screen_library)),
    "Top 10\nfor Docking\nn=10"
  ), levels = rev(c(
    paste0("Full QSAR\nDataset\nn=", nrow(main_df)),
    paste0("Potent Pool\n(pIC50>=7.0)\nn=", n_start),
    paste0("After Ro5\nFilter\nn=", nrow(after_ro5)),
    paste0("After MW/logP\nCleaning\nn=", nrow(after_logp)),
    paste0("Final Screening\nLibrary\nn=", nrow(screen_library)),
    "Top 10\nfor Docking\nn=10"
  ))),
  Count = c(
    nrow(main_df), n_start, nrow(after_ro5),
    nrow(after_logp), nrow(screen_library), 10
  ),
  Fill_Group = c("start","pool","filter","filter","library","selected")
)

fill_vals <- c("start"="#4575b4","pool"="#74add1","filter"="#fee090",
               "library"="#fc8d59","selected"="#d73027")

p_funnel <- ggplot(funnel_data, aes(x=Stage, y=Count, fill=Fill_Group)) +
  geom_col(width=0.7, colour="white", linewidth=0.4) +
  geom_text(aes(label=scales::comma(Count)),
            hjust=-0.12, fontface="bold", size=4.2) +
  coord_flip() +
  scale_fill_manual(values=fill_vals) +
  scale_y_continuous(expand=expansion(mult=c(0, 0.18)),
                     labels=scales::comma) +
  labs(
    title    = "Figure T6-1: Virtual Screening Funnel -- AChE Inhibitor Candidate Selection",
    subtitle = "Sequential filters reduce 5,604 compounds to 10 high-quality docking candidates",
    x        = NULL,
    y        = "Number of Compounds",
    caption  = paste0("Final library: ", nrow(screen_library),
                      " drug-like potent AChE inhibitors | Top 10 selected by composite score.")
  ) +
  theme_bw(base_size=12) +
  theme(
    plot.title      = element_text(face="bold", size=11),
    plot.subtitle   = element_text(size=9, colour="grey40"),
    legend.position = "none",
    axis.text.y     = element_text(size=9)
  )

print(p_funnel)
dir.create("figures", showWarnings=FALSE)
ggsave("figures/Figure_T6_1_Screening_Funnel.png", p_funnel, width=11, height=5, dpi=150)
cat("Saved: figures/Figure_T6_1_Screening_Funnel.png\n")
```

------------------------------------------------------------------------

# Figure T6-2: Composite Score Distribution -- Full Library

```{r fig-score-dist, fig.height=5}
# Cutoff score at top 10
cutoff_score <- ranked_library$Composite_Score[10]

p_score <- ggplot(ranked_library, aes(x=Composite_Score)) +
  geom_histogram(aes(fill=after_stat(x) >= cutoff_score),
                 binwidth=0.005, colour="white", linewidth=0.2) +
  scale_fill_manual(values=c("TRUE"="#d73027","FALSE"="#4575b4"),
                    labels=c("TRUE"="Top 10 selected","FALSE"="Not selected")) +
  geom_vline(xintercept=cutoff_score, linetype="dashed",
             colour="#d73027", linewidth=1) +
  annotate("text", x=cutoff_score+0.002, y=Inf,
           label=paste0("Top-10 cutoff\n= ", round(cutoff_score,4)),
           vjust=1.5, hjust=0, size=3.2, colour="#d73027", fontface="bold") +
  scale_x_continuous(breaks=pretty_breaks(8)) +
  labs(
    title    = "Figure T6-2: Composite Score Distribution -- Virtual Screening Library",
    subtitle = paste0("n = ", nrow(ranked_library), " drug-like AChE inhibitors | ",
                      "Red = top 10 selected for docking"),
    x        = "Composite Score  (pIC50: 50%, LE: 25%, Drug-likeness: 15%, logP: 10%)",
    y        = "Number of Compounds",
    fill     = NULL,
    caption  = "Higher score = better multi-parameter candidate. Cutoff at rank 10."
  ) +
  theme_bw(base_size=12) +
  theme(
    plot.title      = element_text(face="bold", size=11),
    plot.subtitle   = element_text(size=9, colour="grey40"),
    legend.position = "top"
  )

print(p_score)
ggsave("figures/Figure_T6_2_Score_Distribution.png", p_score, width=10, height=5, dpi=150)
cat("Saved: figures/Figure_T6_2_Score_Distribution.png\n")
```

------------------------------------------------------------------------

# Figure T6-3: pIC50 vs Composite Score (Top 10 Highlighted)

```{r fig-pic50-score, fig.height=6}
plot_df <- ranked_library %>%
  mutate(
    Is_Top10  = Rank <= 10,
    Label     = if_else(Is_Top10, ChEMBL_ID, NA_character_),
    Cat       = case_when(
      Is_Top10          ~ "Top 10 (selected for docking)",
      pIC50 >= 9.0      ~ "Extremely Potent (>=9.0)",
      pIC50 >= 8.0      ~ "Very Potent (8.0-9.0)",
      pIC50 >= 7.5      ~ "Potent (7.5-8.0)",
      TRUE              ~ "Moderate (7.0-7.5)"
    )
  )

cat_cols <- c(
  "Top 10 (selected for docking)" = "#d73027",
  "Extremely Potent (>=9.0)"       = "#b15928",
  "Very Potent (8.0-9.0)"         = "#fc8d59",
  "Potent (7.5-8.0)"              = "#4575b4",
  "Moderate (7.0-7.5)"            = "#91bfdb"
)
cat_size <- c("Top 10 (selected for docking)"=3.5,
              "Extremely Potent (>=9.0)"=2,
              "Very Potent (8.0-9.0)"=1.8,
              "Potent (7.5-8.0)"=1.5,
              "Moderate (7.0-7.5)"=1.2)

p_scatter <- ggplot(plot_df %>% arrange(Is_Top10),
                    aes(x=pIC50, y=Composite_Score, colour=Cat, size=Cat)) +
  geom_point(alpha=0.65) +
  geom_label_repel(
    aes(label=Label), size=2.5, max.overlaps=20,
    box.padding=0.5, point.padding=0.4,
    segment.colour="grey60", segment.size=0.3,
    fill="white", colour="#c62828", fontface="bold",
    na.rm=TRUE
  ) +
  scale_colour_manual(values=cat_cols) +
  scale_size_manual(values=cat_size) +
  labs(
    title    = "Figure T6-3: pIC50 vs Composite Score -- Virtual Screening Library",
    subtitle = paste0("n = ", nrow(ranked_library), " compounds | Top 10 labelled in red"),
    x        = "pIC50 (ChEMBL validated experimental value)",
    y        = "Composite Score",
    colour   = "Category", size = "Category",
    caption  = "Top 10 selected compounds cluster at high pIC50 AND high composite score."
  ) +
  theme_bw(base_size=12) +
  theme(
    plot.title      = element_text(face="bold", size=11),
    plot.subtitle   = element_text(size=9, colour="grey40"),
    legend.position = "bottom",
    legend.text     = element_text(size=8)
  )

print(p_scatter)
ggsave("figures/Figure_T6_3_pIC50_vs_Score.png", p_scatter, width=11, height=6, dpi=150)
cat("Saved: figures/Figure_T6_3_pIC50_vs_Score.png\n")
```

------------------------------------------------------------------------

# Figure T6-4: Chemical Space -- MW vs logP

```{r fig-chemical-space, fig.height=6}
p_chem <- ggplot(plot_df %>% arrange(Is_Top10),
                 aes(x=MW, y=logP, colour=Cat, size=Cat)) +
  geom_point(alpha=0.60) +
  geom_label_repel(
    aes(label=Label), size=2.5, max.overlaps=20,
    box.padding=0.5, point.padding=0.4,
    segment.colour="grey60", segment.size=0.3,
    fill="white", colour="#c62828", fontface="bold",
    na.rm=TRUE
  ) +
  # Lipinski Ro5 box
  annotate("rect", xmin=150, xmax=500, ymin=-2, ymax=5,
           fill=NA, colour="grey40", linetype="dashed", linewidth=0.5) +
  annotate("text", x=488, y=4.7, label="Lipinski Ro5 space",
           hjust=1, size=3, colour="grey40") +
  # CNS MPO optimal logP zone
  annotate("rect", xmin=150, xmax=500, ymin=1.0, ymax=3.5,
           fill="#e8f5e9", alpha=0.25) +
  annotate("text", x=155, y=3.3, label="Optimal CNS logP zone (1-3.5)",
           hjust=0, size=2.8, colour="#2e7d32") +
  scale_colour_manual(values=cat_cols) +
  scale_size_manual(values=cat_size) +
  scale_x_continuous(breaks=seq(150,500,50)) +
  labs(
    title    = "Figure T6-4: Chemical Space of Screening Library -- MW vs logP",
    subtitle = "All compounds within Lipinski Ro5 space | CNS optimal logP zone shaded green",
    x        = "Molecular Weight (Da)",
    y        = "logP",
    colour   = "Category", size="Category",
    caption  = paste0("n = ", nrow(ranked_library),
                      " compounds | Dashed box = Lipinski Ro5 boundary")
  ) +
  theme_bw(base_size=12) +
  theme(
    plot.title      = element_text(face="bold", size=11),
    plot.subtitle   = element_text(size=9, colour="grey40"),
    legend.position = "bottom",
    legend.text     = element_text(size=8)
  )

print(p_chem)
ggsave("figures/Figure_T6_4_ChemSpace_MW_logP.png", p_chem, width=11, height=6, dpi=150)
cat("Saved: figures/Figure_T6_4_ChemSpace_MW_logP.png\n")
```

------------------------------------------------------------------------

# Figure T6-5: Ligand Efficiency -- Library vs Top 10

```{r fig-le-ridge, fig.height=5}
le_ridge_df <- bind_rows(
  ranked_library %>% mutate(Group = "Full Screening Library"),
  top10          %>% mutate(Group = "Top 10 Selected")
) %>%
  filter(!is.na(LE_calc))

p_le <- ggplot(le_ridge_df, aes(x=LE_calc, y=Group, fill=Group)) +
  geom_density_ridges(
    alpha=0.75, scale=0.85,
    quantile_lines=TRUE, quantiles=2,
    jittered_points=TRUE,
    point_size=0.9, point_alpha=0.5,
    point_colour="grey30"
  ) +
  scale_fill_manual(values=c("Full Screening Library"="#4575b4",
                              "Top 10 Selected"="#d73027")) +
  labs(
    title    = "Figure T6-5: Ligand Efficiency (LE) Distribution -- Library vs Top 10",
    subtitle = "LE = pIC50 / Heavy Atom Count | Vertical line = median LE",
    x        = "Ligand Efficiency (LE = pIC50 / Heavy Atoms)",
    y        = NULL, fill=NULL,
    caption  = "Top 10 compounds show enriched LE -- potent AND efficient binding per atom."
  ) +
  theme_bw(base_size=12) +
  theme(
    plot.title      = element_text(face="bold", size=11),
    plot.subtitle   = element_text(size=9, colour="grey40"),
    legend.position = "none"
  )

print(p_le)
ggsave("figures/Figure_T6_5_LE_Ridge.png", p_le, width=10, height=4.5, dpi=150)
cat("Saved: figures/Figure_T6_5_LE_Ridge.png\n")
```

------------------------------------------------------------------------

# Figure T6-6: Top 10 Multi-Property Profile

```{r fig-top10-multibar, fig.height=7}
top10_long <- top10 %>%
  mutate(
    Label    = paste0("#", Rank, "  ", ChEMBL_ID),
    Label    = factor(Label, levels = rev(paste0("#", Rank, "  ", ChEMBL_ID)))
  )

# Panel A: pIC50
pA <- ggplot(top10_long, aes(x=Label, y=pIC50, fill=Composite_Score)) +
  geom_col(colour="white") +
  geom_text(aes(label=round(pIC50,2)), hjust=-0.1, size=3.2, fontface="bold") +
  coord_flip() +
  scale_fill_viridis_c(option="D", name="Composite\nScore") +
  scale_y_continuous(expand=expansion(mult=c(0,0.18))) +
  labs(title="A: pIC50", x=NULL, y="pIC50") +
  theme_bw(base_size=10) +
  theme(legend.position="right")

# Panel B: MW
pB <- ggplot(top10_long, aes(x=Label, y=MW, fill=Composite_Score)) +
  geom_col(colour="white") +
  geom_hline(yintercept=500, linetype="dashed", colour="#c62828", linewidth=0.5) +
  coord_flip() +
  scale_fill_viridis_c(option="D") +
  scale_y_continuous(expand=expansion(mult=c(0,0.12))) +
  labs(title="B: MW (Da)  [dashed = Ro5 limit]", x=NULL, y="MW (Da)") +
  theme_bw(base_size=10) +
  theme(legend.position="none", axis.text.y=element_blank())

# Panel C: LE
pC <- ggplot(top10_long, aes(x=Label, y=LE_calc, fill=Composite_Score)) +
  geom_col(colour="white") +
  geom_text(aes(label=round(LE_calc,3)), hjust=-0.1, size=3, fontface="bold") +
  coord_flip() +
  scale_fill_viridis_c(option="D") +
  scale_y_continuous(expand=expansion(mult=c(0,0.22))) +
  labs(title="C: Ligand Efficiency (LE)", x=NULL, y="LE") +
  theme_bw(base_size=10) +
  theme(legend.position="none", axis.text.y=element_blank())

p_multi <- (pA | pB | pC) +
  plot_annotation(
    title    = "Figure T6-6: Multi-Property Profile of Top 10 Docking Candidates",
    subtitle = "Ranked by composite score | Colour intensity = composite score",
    theme    = theme(
      plot.title    = element_text(face="bold", size=11),
      plot.subtitle = element_text(size=9, colour="grey40")
    )
  )

print(p_multi)
ggsave("figures/Figure_T6_6_Top10_MultiProperty.png", p_multi,
       width=13, height=7, dpi=150)
cat("Saved: figures/Figure_T6_6_Top10_MultiProperty.png\n")
```

------------------------------------------------------------------------

# Figure T6-7: Potency Category Distribution of Screening Library

```{r fig-potency-cat, fig.height=5}
pot_counts <- screen_library %>%
  filter(!is.na(Potency_Cat), Potency_Cat != "") %>%
  count(Potency_Cat) %>%
  mutate(
    Potency_Cat = factor(Potency_Cat,
                         levels = c("Potent (7-8)", "Very Potent (8-9)",
                                    "Extremely Potent (>9)"))
  ) %>%
  filter(!is.na(Potency_Cat))

cat_fill <- c("Potent (7-8)"         = "#4575b4",
              "Very Potent (8-9)"    = "#fc8d59",
              "Extremely Potent (>9)"= "#d73027")

p_cat <- ggplot(pot_counts, aes(x=Potency_Cat, y=n, fill=Potency_Cat)) +
  geom_col(colour="white", width=0.6) +
  geom_text(aes(label=scales::comma(n)), vjust=-0.4,
            fontface="bold", size=4.5) +
  scale_fill_manual(values=cat_fill) +
  scale_y_continuous(expand=expansion(mult=c(0,0.14)),
                     labels=scales::comma) +
  labs(
    title    = "Figure T6-7: Potency Distribution -- Final Screening Library",
    subtitle = paste0("n = ", nrow(screen_library),
                      " drug-like AChE inhibitors after Ro5 filtering"),
    x        = "Potency Category",
    y        = "Number of Compounds",
    fill     = NULL,
    caption  = "All compounds pIC50 >= 7.0 (IC50 <= 100 nM)"
  ) +
  theme_bw(base_size=12) +
  theme(
    plot.title      = element_text(face="bold", size=11),
    plot.subtitle   = element_text(size=9, colour="grey40"),
    legend.position = "none",
    axis.text.x     = element_text(size=10)
  )

print(p_cat)
ggsave("figures/Figure_T6_7_PotencyCategory.png", p_cat, width=8, height=5, dpi=150)
cat("Saved: figures/Figure_T6_7_PotencyCategory.png\n")
```

------------------------------------------------------------------------

# Output Files

## Export Top 10 Data Table

```{r export-top10}
dir.create("outputs", showWarnings=FALSE)

# Full data table
top10_export <- top10 %>%
  mutate(
    Docking_ID   = paste0("ACHE_Cand_", sprintf("%02d", Rank), "_", ChEMBL_ID),
    pIC50        = round(pIC50, 2),
    MW           = round(MW, 1),
    logP         = round(logP, 2),
    LE           = round(LE_calc, 3),
    IC50_nM      = round(IC50_nM, 3),
    Composite_Score = round(Composite_Score, 4)
  ) %>%
  select(Rank, Docking_ID, ChEMBL_ID, SMILES, pIC50, IC50_nM, MW,
         logP, RO5_Violations, LE, Composite_Score, Potency_Cat)

write.csv(top10_export, "outputs/Task6_top10_docking_candidates.csv", row.names=FALSE)
cat("Saved: outputs/Task6_top10_docking_candidates.csv\n")
```

## Export SMILES File (for OpenBabel/AutoDock)

```{r export-smiles}
smiles_export <- top10_export %>%
  select(SMILES, Docking_ID)

write.table(smiles_export,
            "outputs/Task6_top10_smiles.smi",
            row.names=FALSE, col.names=FALSE,
            quote=FALSE, sep="\t")
cat("Saved: outputs/Task6_top10_smiles.smi\n")
cat("\nSMILES file contents:\n")
cat(readLines("outputs/Task6_top10_smiles.smi"), sep="\n")
```

## Batch Docking Shell Script

```{r batch-script}
batch <- r'(#!/bin/bash
# ============================================================
# AChE Batch Docking Script -- Top 10 Candidates
# Requires: AutoDock Vina, OpenBabel (obabel)
# Usage:  bash Task6_run_docking.sh
# Output: docking_results/ folder
# NOTE: Top 5 compounds share a methoxyphenyl-sulfamide scaffold.
# In a production pipeline, a Tanimoto diversity filter (cutoff < 0.7)
# would be applied before this step. All 10 are retained per project scope.
# ============================================================

RECEPTOR="ACHE_rank001_prepared.pdbqt"
CONFIG="vina_config.txt"        # from Task 5 outputs
SMILES_FILE="Task6_top10_smiles.smi"
RESULTS_DIR="docking_results"

mkdir -p $RESULTS_DIR

echo "======================================================"
echo " AChE Docking Pipeline -- Top 10 Candidates"
echo " Receptor: $RECEPTOR"
echo " Config:   $CONFIG"
echo "======================================================"

while IFS=$\'\\t\' read -r smiles name; do
    echo ""
    echo "Processing: $name"

    # Step 1: Generate 3D conformer from SMILES (OpenBabel)
    obabel -:"$smiles" --gen3d -O "${RESULTS_DIR}/${name}.sdf" \
           -h --minimize --ff MMFF94 2>/dev/null

    # Step 2: Convert SDF -> PDBQT (adds rotatable bonds)
    obabel "${RESULTS_DIR}/${name}.sdf" \
           -O "${RESULTS_DIR}/${name}.pdbqt" -xh 2>/dev/null

    # Step 3: Dock with AutoDock Vina
    vina --config  $CONFIG \
         --ligand   "${RESULTS_DIR}/${name}.pdbqt" \
         --out      "${RESULTS_DIR}/${name}_poses.pdbqt" \
         --log      "${RESULTS_DIR}/${name}_log.txt"

    echo "  v Done: $name"
done < "$SMILES_FILE"

echo ""
echo "======================================================"
echo " All docking complete. Extracting best scores..."
echo "======================================================"
echo ""
grep -h "^   1 " ${RESULTS_DIR}/*_log.txt 2>/dev/null | \
  awk -v OFS="\\t" "{print \$2, \$3, \$4}" | \
  sort -k2 -n
)'

writeLines(batch, "outputs/Task6_run_docking.sh")
cat("Saved: outputs/Task6_run_docking.sh\n")
```

------------------------------------------------------------------------

# Table T6-4: Handoff Summary

```{r handoff-table}
handoff <- tibble(
  File = c(
    "outputs/Task6_top10_docking_candidates.csv",
    "outputs/Task6_top10_smiles.smi",
    "outputs/Task6_run_docking.sh",
    "Task 5 outputs/vina_config.txt",
    "Task 5 ACHE_rank001.pdb -> prepare -> .pdbqt"
  ),
  Contents = c(
    "Full data for top 10: ChEMBL IDs, pIC50, MW, logP, LE, SMILES, composite score",
    "Tab-delimited SMILES + docking ID -- direct input to OpenBabel",
    "Bash script: SMILES -> 3D -> PDBQT -> AutoDock Vina for all 10 compounds",
    "Box centre X/Y/Z, box size, exhaustiveness=32 -- use as-is",
    "AlphaFold2 rank 1 model -- prepare with AutoDockTools for receptor PDBQT"
  ),
  Output = rep("Docking & ADMET", 5)
)

kable(handoff,
      caption = "Table T6-4: Files Produced by Task 6 for Docking") %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width=TRUE) %>%
  column_spec(1, bold=TRUE, color="#1565c0") %>%
  column_spec(3, bold=TRUE, color="#1a6e3d")
```

------------------------------------------------------------------------

# Task 6 Summary

## Screening Library Statistics

```{r library-stats}
stats_table <- tibble(
  Metric     = c(
    "Starting pool (pIC50 >= 7.0)",
    "After drug-likeness filters",
    "Final screening library",
    "Library reduction",
    "pIC50 range (library)",
    "MW range (library)",
    "logP range (library)",
    "Compounds with 0 RO5 violations",
    "Top 10 composite score range",
    "Top 10 pIC50 range"
  ),
  Value      = c(
    paste0(n_start, " compounds"),
    paste0(nrow(after_logp), " compounds"),
    paste0(nrow(screen_library), " compounds"),
    paste0(round(100*(1 - nrow(screen_library)/n_start), 1), "% removed"),
    paste0(round(min(screen_library$pIC50),2), " - ",
           round(max(screen_library$pIC50),2)),
    paste0(round(min(screen_library$MW),1), " - ",
           round(max(screen_library$MW),1), " Da"),
    paste0(round(min(screen_library$logP),2), " - ",
           round(max(screen_library$logP),2)),
    paste0(sum(screen_library$RO5_Violations == 0, na.rm=TRUE),
           " (", round(100*mean(screen_library$RO5_Violations == 0, na.rm=TRUE),1), "%)"),
    paste0(round(min(top10$Composite_Score),4), " - ",
           round(max(top10$Composite_Score),4)),
    paste0(round(min(top10$pIC50),2), " - ", round(max(top10$pIC50),2))
  )
)

kable(stats_table,
      caption = "Table T6-5: Task 6 -- Virtual Screening Summary Statistics") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width=FALSE) %>%
  column_spec(1, bold=TRUE) %>%
  column_spec(2, bold=TRUE, color="#1a6e3d")
```

## Figures Produced

```{r deliverables}
deliverables <- tibble(
  Figure = c("T6-1","T6-2","T6-3","T6-4","T6-5","T6-6","T6-7"),
  Title  = c(
    "Virtual Screening Funnel",
    "Composite Score Distribution",
    "pIC50 vs Composite Score scatter",
    "Chemical Space: MW vs logP",
    "Ligand Efficiency ridge plot",
    "Top 10 Multi-Property Profile",
    "Potency Category Distribution"
  ),
  File   = paste0("figures/Figure_T6_", 1:7,
                  c("_Screening_Funnel",
                    "_Score_Distribution",
                    "_pIC50_vs_Score",
                    "_ChemSpace_MW_logP",
                    "_LE_Ridge",
                    "_Top10_MultiProperty",
                    "_PotencyCategory"), ".png"),
  Status = rep("Saved", 7)
)

kable(deliverables,
      caption = "Table T6-6: Task 6 Figures Produced") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width=TRUE) %>%
  column_spec(4, bold=TRUE, color="#1a6e3d")
```

------------------------------------------------------------------------

*Task 6: Virtual Screening and Hit Expansion \| BI-424
Pharmacoinformatics* *Target: Human AChE (CHEMBL220, UniProt P22303) \|
Disease: Alzheimer's Disease* *Screening library:
`r nrow(screen_library)` drug-like AChE inhibitors -\> top 10 for
docking*
