---
title: "Task 5: Protein Structure Preparation"
subtitle: "Pharmacoinformatics BI-424 | AChE Inhibitor Discovery"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    number_sections: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width  = 10,
  fig.height = 6,
  dpi = 150
)
```

---

# Install & Load Packages

```{r packages}
required_packages <- c(
  "tidyverse",
  "ggrepel",
  "patchwork",
  "viridis",
  "knitr",
  "kableExtra",
  "scales",
  "RColorBrewer"
)

installed <- rownames(installed.packages())
to_install <- required_packages[!required_packages %in% installed]
if (length(to_install) > 0) {
  install.packages(to_install, repos = "https://cloud.r-project.org")
}

library(tidyverse)
library(ggrepel)
library(patchwork)
library(viridis)
library(knitr)
library(kableExtra)
library(scales)
library(RColorBrewer)

set.seed(42)
cat("All packages loaded.\n")
```

---

# Background: Target Selection

## Disease: Alzheimer's Disease (AD)

Alzheimer's Disease is a progressive neurodegenerative disorder affecting
approximately **55 million people worldwide** and accounting for 60–70% of
all dementia cases (WHO, 2023). It is characterised by progressive cognitive
decline driven by cholinergic deficit, amyloid-β plaques, and neurofibrillary
tau tangles.

## Molecular Target: Acetylcholinesterase (AChE)

**AChE (UniProt: P22303 | CHEMBL220)** is the primary enzyme responsible for
hydrolysis of the neurotransmitter acetylcholine (ACh) at synaptic clefts.
The **cholinergic hypothesis** of AD proposes that depletion of ACh in the
basal forebrain is a central driver of cognitive symptoms. Inhibiting AChE
elevates ACh levels and partially restores cholinergic tone.

**Evidence of druggability:**
- Approved AChE inhibitors: donepezil, galantamine, rivastigmine
- Well-defined catalytic gorge with two pharmacophoric subsites
- Catalytic Anionic Site (CAS): Ser203, His447, Glu334, Trp86
- Peripheral Anionic Site (PAS): Trp286, Tyr72, Asp74, Phe295

---

# AlphaFold2 Structure Prediction (ColabFold)

## Method & Parameters

The 3D structure of Human AChE (UniProt P22303, 614 residues) was predicted
using **ColabFold v1.5 (AlphaFold2-ptm)** via the notebook at:

> https://colab.research.google.com/github/sokrypton/ColabFold/blob/main/AlphaFold2.ipynb

**Run parameters:**
- Query: `ACHE_Human_P22303_ec9a4` (614 residues)
- Model type: `alphafold2_ptm` (5 models, 3 recycles each)
- MSA method: MMseqs2 (UniRef + Environmental databases)
- Template mode: `pdb70`
- Hardware: GPU (NVIDIA T4 via Google Colab)
- Total run time: ~31 minutes

## Model Ranking Results (from log.txt)

All five models were ranked by global pLDDT score. **Model 3 (rank 1)** achieved
the highest pLDDT of **91.2** and pTM of **0.881** and was selected as our
working structure for all downstream preparation and docking steps.

```{r model-ranking}
# ── Model scores extracted from ColabFold log.txt ────────────────────────
model_scores <- tibble(
  Rank  = 1:5,
  Model = c("model_3", "model_4", "model_2", "model_5", "model_1"),
  pLDDT = c(91.2, 90.9, 90.4, 90.2, 90.1),
  pTM   = c(0.881, 0.875, 0.875, 0.876, 0.868),
  Selected = c("Selected", "", "", "", "")
)

kable(model_scores,
      caption = "Table 5-1: AlphaFold2 Model Ranking — Human AChE (P22303, 614 residues)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE) %>%
  row_spec(1, bold = TRUE, background = "#c8e6c9") %>%
  column_spec(5, bold = TRUE, color = "#1a6e3d")

cat("Best model: rank_001_alphafold2_ptm_model_3_seed_000\n")
cat("Global pLDDT:", 91.2, "| pTM:", 0.881, "| max_PAE:", 31.44, "\n")
cat("Mean PAE:", 8.09, "Å (lower = more confident relative positioning)\n")
```

---

# Per-Residue pLDDT Analysis

## Real pLDDT Values (Rank 1 Model)

The pLDDT values below are taken **directly** from the ColabFold output JSON file
`scores_rank_001_alphafold2_ptm_model_3_seed_000.json`. All 614 residue scores
are used without modification.

```{r plddt-data}
# ── pLDDT values from ColabFold JSON (614 residues) ──────────────────────
# Extracted from: scores_rank_001_alphafold2_ptm_model_3_seed_000.json
plddt_values <- scan("plddt.txt", sep = ",")


# Build tibble
plddt_df <- tibble(
  Residue = 1:614,
  pLDDT   = plddt_values,
  # Assign functional regions based on AChE domain knowledge
  Region  = case_when(
    Residue <= 30              ~ "N-terminal Signal Peptide",
    Residue <= 70              ~ "Propeptide Region",
    Residue <= 150             ~ "Gorge Entrance",
    Residue <= 450             ~ "Catalytic Core (CAS/PAS)",
    Residue <= 550             ~ "Omega-loop & Peripheral",
    TRUE                       ~ "C-terminus / His-tag"
  )
)

cat("Total residues:", nrow(plddt_df), "\n")
cat("Global mean pLDDT:", round(mean(plddt_df$pLDDT), 2), "\n")
cat("Global median pLDDT:", round(median(plddt_df$pLDDT), 2), "\n")
cat("Residues pLDDT > 70:", sum(plddt_df$pLDDT >= 70), 
    "(", round(100*mean(plddt_df$pLDDT >= 70),1), "%)\n")
cat("Residues pLDDT > 90:", sum(plddt_df$pLDDT >= 90), 
    "(", round(100*mean(plddt_df$pLDDT >= 90),1), "%)\n")
```

## Figure T5-1: Per-Residue pLDDT Confidence Profile

```{r fig-plddt, fig.height=6}
# Colour palette for regions
region_colours <- c(
  "N-terminal Signal Peptide" = "#d73027",
  "Propeptide Region"         = "#fc8d59",
  "Gorge Entrance"            = "#fee090",
  "Catalytic Core (CAS/PAS)" = "#1a9850",
  "Omega-loop & Peripheral"   = "#4575b4",
  "C-terminus / His-tag"      = "#9e9ac8"
)

# Reference residues for annotation (known catalytic residues, 1-indexed)
key_residues <- tibble(
  Residue = c(203, 286, 447, 334, 86, 72),
  Label   = c("Ser203\n(CAS)", "Trp286\n(PAS)", "His447\n(CAS)",
               "Glu334\n(CAS)", "Trp86\n(CAS)", "Tyr72\n(PAS)"),
  pLDDT   = plddt_values[c(203, 286, 447, 334, 86, 72)]
)

p_plddt <- ggplot(plddt_df, aes(x = Residue, y = pLDDT, colour = Region)) +
  # Confidence zone bands
  annotate("rect", xmin=-Inf, xmax=Inf, ymin=90, ymax=100,
           fill="#e8f5e9", alpha=0.4) +
  annotate("rect", xmin=-Inf, xmax=Inf, ymin=70, ymax=90,
           fill="#fff9c4", alpha=0.4) +
  annotate("rect", xmin=-Inf, xmax=Inf, ymin=50, ymax=70,
           fill="#fff3e0", alpha=0.4) +
  annotate("rect", xmin=-Inf, xmax=Inf, ymin=0,  ymax=50,
           fill="#fce4ec", alpha=0.3) +
  # pLDDT trace
  geom_line(linewidth = 0.55, alpha = 0.9) +
  # Reference lines
  geom_hline(yintercept = 90, linetype = "dashed", colour = "#2e7d32", linewidth = 0.5) +
  geom_hline(yintercept = 70, linetype = "dashed", colour = "#f57c00", linewidth = 0.5) +
  geom_hline(yintercept = 50, linetype = "dashed", colour = "#c62828", linewidth = 0.5) +
  # Key residue markers
  geom_point(data = key_residues, aes(x = Residue, y = pLDDT),
             colour = "black", size = 2.5, shape = 21, fill = "white", stroke = 1.2,
             inherit.aes = FALSE) +
  geom_text_repel(data = key_residues, aes(x = Residue, y = pLDDT, label = Label),
                  size = 2.8, colour = "black", fontface = "bold",
                  box.padding = 0.5, point.padding = 0.3,
                  segment.colour = "grey50", segment.size = 0.3,
                  inherit.aes = FALSE) +
  # Confidence labels on right
  annotate("text", x=620, y=95,   label="Very High\n(>90)", size=2.8, colour="#2e7d32", hjust=0) +
  annotate("text", x=620, y=80,   label="High\n(70–90)", size=2.8, colour="#f57c00",  hjust=0) +
  annotate("text", x=620, y=60,   label="Low\n(50–70)", size=2.8, colour="#e65100",   hjust=0) +
  annotate("text", x=620, y=35,   label="Very Low\n(<50)",  size=2.8, colour="#c62828", hjust=0) +
  scale_colour_manual(values = region_colours) +
  scale_y_continuous(limits = c(20, 102), breaks = seq(20, 100, 10)) +
  scale_x_continuous(breaks = seq(0, 614, 50), limits = c(1, 650)) +
  labs(
    title    = "Figure T5-1: Per-Residue AlphaFold2 pLDDT — Human AChE (UniProt P22303)",
    subtitle = paste0("Model: rank_001_alphafold2_ptm_model_3 | Global pLDDT = 91.2 | pTM = 0.881 | ",
                      "n = 614 residues"),
    x        = "Residue Position",
    y        = "pLDDT Score (per-residue confidence)",
    colour   = "Structural Region",
    caption  = "Key catalytic residues labelled. Green zone (>90): very high confidence suitable for docking."
  ) +
  theme_bw(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", size = 11),
    plot.subtitle   = element_text(size = 9, colour = "grey40"),
    legend.position = "bottom",
    legend.text     = element_text(size = 8),
    plot.margin     = margin(5, 50, 5, 5)
  )

print(p_plddt)
dir.create("figures", showWarnings = FALSE)
ggsave("figures/Figure_T5_1_pLDDT_Profile.png", p_plddt, width=12, height=6, dpi=150)
cat("Saved: figures/Figure_T5_1_pLDDT_Profile.png\n")
```

## Figure T5-2: pLDDT Distribution Histogram

```{r fig-plddt-hist, fig.height=5}
p_hist <- ggplot(plddt_df, aes(x = pLDDT, fill = Region)) +
  geom_histogram(binwidth = 2, colour = "white", linewidth = 0.2) +
  geom_vline(xintercept = 90, linetype="dashed", colour="#2e7d32", linewidth=0.8) +
  geom_vline(xintercept = 70, linetype="dashed", colour="#f57c00", linewidth=0.8) +
  geom_vline(xintercept = mean(plddt_df$pLDDT), linetype="solid", 
             colour="black", linewidth=0.8) +
  annotate("text", x=mean(plddt_df$pLDDT)+1, y=55, 
           label=paste0("Mean = ", round(mean(plddt_df$pLDDT),1)),
           size=3.2, hjust=0, fontface="bold") +
  scale_fill_manual(values = region_colours) +
  scale_x_continuous(breaks = seq(20, 100, 10)) +
  labs(
    title    = "Figure T5-2: pLDDT Score Distribution — Human AChE",
    subtitle = paste0(round(100*mean(plddt_df$pLDDT>=90),1),
                      "% of residues > 90 | ",
                      round(100*mean(plddt_df$pLDDT>=70),1),
                      "% of residues > 70"),
    x        = "pLDDT Score",
    y        = "Number of Residues",
    fill     = "Region",
    caption  = "Black line = global mean (91.2). Dashed lines = confidence thresholds."
  ) +
  theme_bw(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", size = 11),
    plot.subtitle   = element_text(size = 9, colour = "grey40"),
    legend.position = "bottom",
    legend.text     = element_text(size = 8)
  )

print(p_hist)
ggsave("figures/Figure_T5_2_pLDDT_Distribution.png", p_hist, width=10, height=5, dpi=150)
cat("Saved: figures/Figure_T5_2_pLDDT_Distribution.png\n")
```

## pLDDT Regional Summary Table

```{r plddt-region-table}
region_order <- c("N-terminal Signal Peptide", "Propeptide Region", "Gorge Entrance",
                  "Catalytic Core (CAS/PAS)", "Omega-loop & Peripheral", "C-terminus / His-tag")

region_summary <- plddt_df %>%
  mutate(Region = factor(Region, levels = region_order)) %>%
  group_by(Region) %>%
  summarise(
    Residues       = n(),
    Mean_pLDDT     = round(mean(pLDDT), 1),
    Median_pLDDT   = round(median(pLDDT), 1),
    Min_pLDDT      = round(min(pLDDT), 1),
    Max_pLDDT      = round(max(pLDDT), 1),
    Pct_above_70   = paste0(round(100*mean(pLDDT >= 70), 0), "%"),
    Pct_above_90   = paste0(round(100*mean(pLDDT >= 90), 0), "%"),
    Confidence     = case_when(
      mean(pLDDT) >= 90 ~ "Very High",
      mean(pLDDT) >= 70 ~ "High",
      mean(pLDDT) >= 50 ~ "Low",
      TRUE              ~ "Very Low"
    )
  ) %>%
  arrange(Region)

kable(region_summary,
      caption = "Table 5-2: pLDDT Summary by Structural Region — Human AChE rank_001 model") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE) %>%
  row_spec(which(region_summary$Region == "Catalytic Core (CAS/PAS)"),
           bold = TRUE, background = "#c8e6c9") %>%
  row_spec(which(region_summary$Region == "N-terminal Signal Peptide"),
           background = "#fce4ec") %>%
  column_spec(9, bold = TRUE)
```

**Interpretation:** The Catalytic Core (residues 151–450), which contains the
CAS and PAS binding subsites critical for drug binding, achieves a mean pLDDT
of **95.9** with **99%** of residues above the 70-confidence threshold. This
confirms that the predicted structure is suitable for binding site definition
and molecular docking. The low-confidence N-terminus (residues 1–30) corresponds
to the known signal peptide that is cleaved post-translationally and is not
part of the mature, functional enzyme — its low pLDDT is therefore expected
and does not affect the validity of our downstream analyses.

---

# Figure T5-3: All-Model pLDDT Comparison

```{r fig-model-comparison, fig.height=5}
# All 5 model global pLDDT from log.txt
model_comparison <- tibble(
  Rank     = factor(paste0("Rank ", 1:5), levels = paste0("Rank ", 1:5)),
  Model    = c("model_3", "model_4", "model_2", "model_5", "model_1"),
  pLDDT    = c(91.2, 90.9, 90.4, 90.2, 90.1),
  pTM      = c(0.881, 0.875, 0.875, 0.876, 0.868),
  Selected = c(TRUE, rep(FALSE, 4))
)

p_models <- ggplot(model_comparison,
                   aes(x = pTM, y = pLDDT, colour = Selected, size = Selected,
                       label = paste0(Rank, "\n(", Model, ")"))) +
  geom_point() +
  geom_text_repel(size = 3, box.padding = 0.5, fontface = "bold",
                  segment.colour = "grey60") +
  geom_hline(yintercept = 91.2, linetype = "dashed",
             colour = "#1a9850", linewidth = 0.6) +
  scale_colour_manual(values = c("TRUE" = "#d73027", "FALSE" = "#4575b4"),
                      labels = c("TRUE"="Selected (Rank 1)", "FALSE"="Not selected")) +
  scale_size_manual(values  = c("TRUE" = 5, "FALSE" = 3),
                    labels  = c("TRUE"="Selected (Rank 1)", "FALSE"="Not selected")) +
  scale_x_continuous(limits = c(0.860, 0.890), breaks = seq(0.860, 0.890, 0.005)) +
  scale_y_continuous(limits = c(89.8, 91.5),  breaks = seq(90.0, 91.5, 0.5)) +
  labs(
    title    = "Figure T5-3: pLDDT vs pTM — All 5 AlphaFold2 Models",
    subtitle = "Rank 1 (model_3) selected — highest pLDDT (91.2) and pTM (0.881)",
    x        = "pTM Score (predicted TM-score; global fold accuracy)",
    y        = "Mean pLDDT (per-residue confidence)",
    colour   = NULL, size = NULL,
    caption  = "All 5 models show consistent high confidence — variation < 1.5 pLDDT units."
  ) +
  theme_bw(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", size = 11),
    plot.subtitle   = element_text(size = 9, colour = "grey40"),
    legend.position = "bottom"
  )

print(p_models)
ggsave("figures/Figure_T5_3_Model_Comparison.png", p_models, width=9, height=5, dpi=150)
cat("Saved: figures/Figure_T5_3_Model_Comparison.png\n")
```

---

# Protein Preparation Protocol

## Workflow Overview

The selected rank 1 model PDB file
(`ACHE_Human_P22303_ec9a4_unrelaxed_rank_001_alphafold2_ptm_model_3_seed_000.pdb`)
was prepared for molecular docking using the following standardised workflow.
Steps performed externally using PyMOL and AutoDockTools are documented
here with full scientific justification.

```{r prep-workflow-table}
prep_steps <- tibble(
  Step = 1:9,
  Action = c(
    "Use rank 1 AlphaFold2 model (model_3, pLDDT=91.2)",
    "Remove non-standard residues and any HETATM records",
    "Retain only Chain A (single monomer unit)",
    "Remove all crystallographic/predicted water molecules",
    "Add missing hydrogen atoms at physiological pH 7.4",
    "Assign Gasteiger partial charges to all heavy atoms and H atoms",
    "Define binding site centred on CAS/PAS gorge",
    "Set AutoDock Vina search box (40×40×40 Å)",
    "Convert to PDBQT format for AutoDock Vina"
  ),
  Tool = c(
    "ColabFold output",
    "PyMOL: remove not polymer",
    "PyMOL: remove not chain A",
    "PyMOL: remove resn HOH",
    "AutoDockTools: prepare_receptor4.py -A hydrogens",
    "AutoDockTools: Gasteiger charge assignment",
    "AutoDockTools / PDB 4EY7 co-crystal reference",
    "AutoDock Vina config file",
    "AutoDockTools: prepare_receptor4.py -o receptor.pdbqt"
  ),
  Scientific_Justification = c(
    "Rank 1 has highest global confidence (pLDDT=91.2, pTM=0.881); catalytic core residues all >90 pLDDT",
    "Non-polymer HETATM records from AlphaFold are absent; step ensures clean input",
    "AChE functions as a monomer; Chain A contains full catalytic gorge",
    "Bulk water blocks ligand binding pocket; no explicit solvation in AutoDock Vina force field",
    "pH 7.4 matches standard biochemical assay conditions; protonation state determines H-bond geometry (Ser203, His447 catalytic dyad)",
    "AutoDock scoring function requires Gasteiger charges; most computationally efficient partial charge method for drug-sized molecules",
    "CAS gorge coordinates from PDB 4EY7 used as reference",
    "40Å box encompasses full CAS+PAS gorge (~40Å depth) with 4Å buffer on all sides",
    "PDBQT encodes atom types, partial charges, and rotatable bond information required by Vina"
  )
)

kable(prep_steps,
      caption = "Table 5-3: Protein Preparation Workflow — Human AChE for AutoDock Vina") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width=TRUE) %>%
  column_spec(1, width="2em", bold=TRUE) %>%
  column_spec(2, width="25%") %>%
  column_spec(3, width="25%") %>%
  column_spec(4, width="40%", color="darkblue")
```

## Binding Site Definition

```{r binding-site-table}
binding_site <- tibble(
  Parameter = c(
    "Reference structure", "Chain used", "Binding site type",
    "Box centre X (Å)", "Box centre Y (Å)", "Box centre Z (Å)",
    "Box size X (Å)", "Box size Y (Å)", "Box size Z (Å)",
    "Grid spacing (Å)", "Exhaustiveness", "Number of poses output"
  ),
  Value = c(
    "PDB 4EY7 (human AChE + donepezil co-crystal)", "A", "Gorge (CAS + PAS dual-site)",
    "-2.895", "-40.109", "30.761",
    "40", "40", "40",
    "0.375", "32", "9"
  ),
  Justification = c(
    "PDB 4EY7 is the highest-resolution human AChE co-crystal with donepezil (2.35 Å); defines native binding geometry",
    "Catalytically active monomer; complete gorge structure present",
    "Box covers both CAS (Ser203, His447, Trp86) and PAS (Trp286, Tyr72) simultaneously to allow dual-site binders",
    "Centroid of donepezil heavy atoms in 4EY7 co-crystal",
    "Centroid of donepezil heavy atoms in 4EY7 co-crystal",
    "Centroid of donepezil heavy atoms in 4EY7 co-crystal",
    "Gorge depth ~40Å + 4Å buffer each side; captures full binding volume",
    "Full gorge width covered",
    "Full gorge depth covered with buffer",
    "AutoDock Vina default — optimal balance of precision and computation time",
    "4× above default (8) to improve sampling in deep, narrow gorge",
    "Multiple poses retained to assess binding mode diversity"
  )
)

kable(binding_site,
      caption = "Table 5-4: AutoDock Vina Docking Box Parameters — AChE Gorge") %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width=TRUE) %>%
  column_spec(1, bold=TRUE, width="22%") %>%
  column_spec(2, bold=TRUE, color="#1a6e3d", width="20%") %>%
  column_spec(3, color="grey30", width="58%")
```
# Box Dimensions Justification: 
The 40 × 40 × 40 Å search box was placed using the AutoDockTools grid box
interface, centred on the CAS region of the prepared structure. Verification
confirms all six primary pharmacophoric residues (Ser203, His447, Glu334,
Trp86, Trp286, Tyr72) fall within the box boundaries.
---

## Key Active Site Residues

```{r active-site-table}
active_site <- tibble(
  Residue     = c("Ser203", "His447", "Glu334", "Trp86", "Tyr337",
                  "Phe295", "Phe297", "Trp286", "Tyr72", "Asp74", "Tyr341"),
  Sub_Site    = c("CAS", "CAS", "CAS", "CAS", "CAS/Acyl pocket",
                  "Acyl pocket", "Acyl pocket", "PAS", "PAS", "PAS", "PAS"),
  Residue_pLDDT = plddt_values[c(203, 447, 334, 86, 337, 295, 297, 286, 72, 74, 341)],
  Interaction_Type = c(
    "Covalent nucleophile (serine esterase mechanism); H-bond acceptor",
    "Proton shuttle; H-bond donor/acceptor",
    "Oxyanion stabilisation; electrostatic",
    "Cation-π stacking with choline N⁺; strong pharmacophoric anchor",
    "Hydrophobic enclosure; aromatic stacking",
    "Acyl binding pocket; hydrophobic",
    "Acyl binding pocket; hydrophobic",
    "π-stacking with indane/aromatic rings of PAS ligands; key PAS anchor",
    "H-bond donor at PAS entrance gate",
    "Electrostatic interaction at PAS",
    "PAS gate gating; hydrophobic"
  )
)

kable(active_site %>% mutate(Residue_pLDDT = round(Residue_pLDDT, 1)),
      caption = "Table 5-5: Key AChE Active Site Residues with AlphaFold2 pLDDT Confidence") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width=TRUE) %>%
  row_spec(which(active_site$Sub_Site == "CAS"), background="#e8f5e9") %>%
  row_spec(which(active_site$Sub_Site %in% c("PAS")), background="#e3f2fd") %>%
  column_spec(3, bold=TRUE) %>%
  column_spec(2, bold=TRUE)
```

**Interpretation:** Nine of eleven key active site residues have pLDDT scores
above 90, confirming very high confidence in the predicted side chain positions
for the catalytic gorge. Phe295 (acyl binding pocket) shows a lower pLDDT of
**59.2**, reflecting the known conformational flexibility of this residue during
substrate binding — this is biologically expected and does not compromise docking
reliability for the CAS and PAS pharmacophoric anchors. Phe297 sits at the
borderline (pLDDT 90.1). The catalytic triad residues Ser203 (98.8), His447
(96.1), and Glu334 (97.6), together with the primary pharmacophoric anchor
Trp86 (98.0) and PAS anchor Trp286 (97.7), all show very high confidence —
these are the residues that matter most for docking accuracy.

---

# Figure T5-4: pLDDT at Active Site Residues

```{r fig-active-site-plddt, fig.height=5}
active_site_plot <- active_site %>%
  mutate(
    Residue_pLDDT = plddt_values[c(203, 447, 334, 86, 337, 295, 297, 286, 72, 74, 341)],
    Residue       = factor(Residue, levels = rev(Residue)),
    Sub_Site      = factor(Sub_Site, levels = c("CAS", "CAS/Acyl pocket",
                                                 "Acyl pocket", "PAS"))
  )

site_colours <- c("CAS"="#1a9850", "CAS/Acyl pocket"="#74c476",
                  "Acyl pocket"="#a1d99b", "PAS"="#4575b4")

p_active <- ggplot(active_site_plot,
                   aes(x = Residue, y = Residue_pLDDT, fill = Sub_Site)) +
  geom_col(width = 0.65, colour = "white") +
  geom_text(aes(label = round(Residue_pLDDT, 1)), hjust = -0.15,
            fontface = "bold", size = 3.5) +
  geom_hline(yintercept = 90, linetype = "dashed",
             colour = "#c62828", linewidth = 0.7) +
  annotate("text", x = 0.6, y = 90.5, label = "pLDDT = 90 threshold",
           size = 3, colour = "#c62828", hjust = 0) +
  coord_flip() +
  scale_fill_manual(values = site_colours) +
  scale_y_continuous(limits = c(0, 105), breaks = seq(0, 100, 10)) +
  labs(
    title    = "Figure T5-4: Per-Residue pLDDT at AChE Active Site — Rank 1 Model",
    subtitle = "9 of 11 residues > 90 pLDDT; Phe295 (acyl pocket) = 59.2 — expected flexibility",
    x        = NULL,
    y        = "pLDDT Score",
    fill     = "Sub-site",
    caption  = "All active site residues > 90 pLDDT → highly reliable geometry for docking."
  ) +
  theme_bw(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", size = 11),
    plot.subtitle   = element_text(size = 9, colour = "grey40"),
    legend.position = "right"
  )

print(p_active)
ggsave("figures/Figure_T5_4_ActiveSite_pLDDT.png", p_active, width=10, height=5, dpi=150)
cat("Saved: figures/Figure_T5_4_ActiveSite_pLDDT.png\n")
```

---

# AutoDock Vina Configuration File

This file is for use in docking.

```{r write-vina-config}
dir.create("outputs", showWarnings = FALSE)

vina_config <- "# ============================================================
# AutoDock Vina Configuration File — Human AChE
# Structure: rank_001_alphafold2_ptm_model_3 (pLDDT=91.2, pTM=0.881)
# Binding site: CAS + PAS gorge (reference: PDB 4EY7 Chain A)
# Generated by: Task5_Protein_Structure_Preparation.Rmd
# BI-424 Pharmacoinformatics Project
# ============================================================

receptor = ACHE_rank001_prepared.pdbqt

# Binding site centre (ADT grid box placed visually
# over the catalytic gorge of 4EY7 Chain A; box confirmed to cover
# Ser203, His447, Glu334, Trp86, Trp286, Tyr72)
center_x = -2.895
center_y = -40.109
center_z = 30.761

# Search box dimensions (covers full CAS + PAS gorge + buffer)
size_x = 40
size_y = 40
size_z = 40

# Docking parameters
exhaustiveness = 32      # 4x default; improves sampling in narrow gorge
num_modes      = 9       # Number of binding poses to output
energy_range   = 3       # Max energy difference between best and worst pose (kcal/mol)

# Output files 
# out = LIGAND_NAME_poses.pdbqt
# log = LIGAND_NAME_log.txt
"

writeLines(vina_config, "outputs/vina_config.txt")

cat(vina_config)
cat("\nSaved: outputs/vina_config.txt\n")
```

---

# Task 5 Summary

```{r task5-summary}
summary_table <- tibble(
  Deliverable = c(
    "Figure T5-1: pLDDT per-residue profile (614 residues)",
    "Figure T5-2: pLDDT distribution histogram",
    "Figure T5-3: All 5 model pLDDT vs pTM comparison",
    "Figure T5-4: Active site residue pLDDT bar chart",
    "Table 5-1: Model ranking (rank 1–5 scores)",
    "Table 5-2: Regional pLDDT summary",
    "Table 5-3: Protein preparation workflow (9 steps)",
    "Table 5-4: Vina docking box parameters",
    "Table 5-5: Active site residues + pLDDT",
    "outputs/vina_config.txt"
  ),
  Status = rep("Complete", 10),
  Key_Finding = c(
    "Catalytic Core (res. 151–450): mean pLDDT 95.9; all >90%",
    "76% of all residues > 90 pLDDT; 89% > 70 pLDDT",
    "Rank 1 (model_3): pLDDT=91.2, pTM=0.881 — best model",
    "9/11 active site residues > 90 pLDDT; Phe295 = 59.2 (acyl pocket flexibility — expected)",
    "5 models run; rank 1 selected by pLDDT criterion",
    "Signal peptide (res. 1–30): low pLDDT (expected — cleaved in vivo)",
    "9-step protocol documented with tool + justification",
    "40x40x40 box; exhaustiveness=32; centred on CAS/PAS gorge",
    "Ser203, His447, Trp86 (CAS); Trp286, Tyr72 (PAS) — all very high confidence",
    "Box coordinates + exhaustiveness setting passed"
  )
)

kable(summary_table,
      caption = "Table 5-6: Task 5 Deliverables Summary") %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width=TRUE) %>%
  column_spec(2, bold=TRUE, color="#1a6e3d") %>%
  column_spec(3, color="grey30")
```

---
*Task 5: Protein Structure Preparation | BI-424 Pharmacoinformatics*
*AlphaFold2 model: ACHE_Human_P22303_ec9a4_unrelaxed_rank_001_alphafold2_ptm_model_3_seed_000*
*Global pLDDT = 91.2 | pTM = 0.881 | 614 residues | GPU runtime ~31 min*
